### 2004-06-28 Documentation/Manual.tex
*** empty log message ***
two files added for the GPL and FDL
latest version of part2

\usepackage{cite}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{a4wide}
\usepackage{fancyheadings}
\usepackage[footnotesize,bf]{caption}
\usepackage{ae,aecompl} % only for Type 1 fonts

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\ra}{$\rightarrow$}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\nn}{\nonumber}
\newcommand{\ldm}{$\Delta m_{31}^2$}
\newcommand{\sdm}{$\Delta m_{21}^2$}
\newcommand{\deltacp}{$\delta_{\mathrm{CP}}$}
\newcommand{\stheta}{$\sin^2 2 \theta_{13}$}

\newcommand{\ie}{{\it i.e.}}
\newcommand{\Ie}{{\it I.e.}}
\newcommand{\eg}{{\it e.g.}}
\newcommand{\Eg}{{\it E.g.}}
\newcommand{\cf}{{\it cf.}}
\newcommand{\etc}{{\it etc.}}
\newcommand{\eq}{Eq.}
\newcommand{\eqs}{Eqs.}
\newcommand{\Def}{Definition}
\newcommand{\fig}{Fig.}
\newcommand{\Fig}{Fig.}
\newcommand{\figs}{Figs.}
\newcommand{\Figs}{Figs.}
\newcommand{\Ref}{Ref.}
\newcommand{\Refs}{Refs.}
\newcommand{\Sec}{Sec.}
\newcommand{\Secs}{Secs.}
\newcommand{\App}{the Appendix}
\newcommand{\Apps}{Appendices}
\newcommand{\Tab}{Table}
\newcommand{\Tabs}{Tables}

\newcommand{\JHFSK}{{\sc JHF-SK}}
\newcommand{\NUMI }{{\sc NuMI}}
\newcommand{\ReactorI}{{\sc Reactor-I}}
\newcommand{\ReactorII}{{\sc Reactor-II}}
\newcommand{\JHFHK}{{\sc JHF-HK}}
\newcommand{\NuFactI}{{\sc NuFact-I}}
\newcommand{\NuFactII}{{\sc NuFact-II}}

\newcommand{\GLOBES}{{\sf GLoBES}}
\newcommand{\EDM}{{\sf EDM}}

\newcommand{\equ}[1]{\eq~(\ref{equ:#1})}
\newcommand{\figu}[1]{\fig~\ref{fig:#1}}

@@

\begin{center}
{ \Large \bf
\GLOBES\ 1.X \\
Global Long Baseline Experiment Simulator \\ }
\vspace*{0.5cm}
{\large \bf User's and reference manual }
\end{center}

@@

\begin{center}
Here comes our logo!
\end{center}

@@

\chapter*{What is \GLOBES\ good for?}

Here comes a maximum of one page what \GLOBES\ is good for and what main features it has.
\cleardoublepage
\tableofcontents

@@

@@

>>>>

\usepackage{verbatim}
\usepackage{cite}
\usepackage[dvips]{color}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{a4wide}
\usepackage{fancyheadings}
\usepackage[footnotesize,bf]{caption}
\usepackage{makeidx}
\usepackage{ae,aecompl} % only for Type 1 fonts

\newcommand{\capdef}{}
\newcommand{\mycaption}[2][\capdef]{\renewcommand{\capdef}{#2}%
        \caption[#1]{{ %\itshape
 {\footnotesize #2}}}}
\makeatletter
\renewcommand{\fnum@table}{\textbf{\tablename~\thetable}}
\renewcommand{\fnum@figure}{\textbf{\figurename~\thefigure}}
\makeatother

% Index: use "makeindex Manual" to create .ind file
\makeindex

\newcommand{\bi}{\begin{itemize}}
\newcommand{\ei}{\end{itemize}}
\newcommand{\ra}{$\rightarrow$}
\newcommand{\be}{\begin{equation}}
\newcommand{\ee}{\end{equation}}
\newcommand{\bea}{\begin{eqnarray}}
\newcommand{\eea}{\end{eqnarray}}
\newcommand{\nn}{\nonumber}
\newcommand{\ldm}{\Delta m_{31}^2}
\newcommand{\sdm}{\Delta m_{21}^2}
\newcommand{\deltacp}{\delta_{\mathrm{CP}}}
\newcommand{\stheta}{\sin^2 2 \theta_{13}}

\newcommand{\ie}{{\it i.e.}}
\newcommand{\Ie}{{\it I.e.}}
\newcommand{\eg}{{\it e.g.}}
\newcommand{\Eg}{{\it E.g.}}
\newcommand{\cf}{{\it cf.}}
\newcommand{\etc}{{\it etc.}}
\newcommand{\eq}{Eq.}
\newcommand{\eqs}{Eqs.}
\newcommand{\Def}{Definition}
\newcommand{\fig}{Fig.}
\newcommand{\Fig}{Fig.}
\newcommand{\figs}{Figs.}
\newcommand{\Figs}{Figs.}
\newcommand{\Ref}{Ref.}
\newcommand{\Refs}{Refs.}
\newcommand{\Sec}{Sec.}
\newcommand{\Secs}{Secs.}
\newcommand{\Chapt}{Chapter}
\newcommand{\Chapts}{Chapters}
\newcommand{\Part}{Part}
\newcommand{\App}{the Appendix}
\newcommand{\Apps}{Appendices}
\newcommand{\Tab}{Table}
\newcommand{\Tabs}{Tables}

\newcommand{\JHFSK}{{\sc JHF-SK}}
\newcommand{\NUMI }{{\sc NuMI}}
\newcommand{\ReactorI}{{\sc Reactor-I}}
\newcommand{\ReactorII}{{\sc Reactor-II}}
\newcommand{\JHFHK}{{\sc JHF-HK}}
\newcommand{\NuFactI}{{\sc NuFact-I}}
\newcommand{\NuFactII}{{\sc NuFact-II}}
\newcommand{\SPL}{{\sc SPL}}
\newcommand{\Beta}{{\sc $\beta$-Beam}}

\newcommand{\GLOBES}{{\sf GLoBES}}
\newcommand{\AEDL}{{\sf AEDL}}
\newcommand{\EDM}{{\sf EDM}}

\newcommand{\equ}[1]{\eq~(\ref{equ:#1})}
\newcommand{\figu}[1]{\fig~\ref{fig:#1}}
\newcommand{\tabl}[1]{\Tab~\ref{tab:#1}}
\newcommand{\tb}{\hspace*{3ex}}

\definecolor{light}{gray}{0.93}
\definecolor{heavy}{gray}{0.35}
\setlength{\fboxrule}{0.1cm}
\newcommand{\example}[2]{
\begin{table}[p]
\begin{center}
\fcolorbox{heavy}{light}{
\begin{minipage}[t][\textheight][t]{15.5cm}
{\em \underline{Example:} #1}

\vspace*{0.3cm}

#2

\end{minipage}
}
\end{center}
\end{table}
}

\newtheorem{function}{Function}[chapter]

@@

\begin{center}
{ \Large \bf
\GLOBES\ 1.X \\
General Long Baseline Experiment Simulator \\ }
\vspace*{0.5cm}
{\large \bf User's and reference manual }
\end{center}

@@

\begin{center}
\colorbox{black}{\includegraphics[width=7cm]{earthint}}

\vspace*{-1cm}

{\Huge \textcolor{yellow}{GLoBES}}
\end{center}

@@

\chapter*{What is \GLOBES ?}

\GLOBES\ (``General Long Baseline Experiment Simulator'') is a flexible
software package to simulate neutrino long baseline experiments. On the
one hand, it contains a comprehensive abstract experiment definition
language, which allows to describe most classes of long baseline experinents
at an abstract level. On the other hand, it provides a C-library to 
process the experiment information in order to obtain oscillation
probabilities, rate vectors, and $\Delta \chi^2$-values. Currently, 
\GLOBES\ is available for Linux. Since the source code is included,
the modifications to other operating systems should be doable.

\GLOBES\ allows to simulate experiments with stationary neutrino point sources, where each experiment is assumed to have only one neutrino source.
Such experiments are neutrino beam experiments and reactor experiments. 
Geometrical effects of a source distribution, such as in the sun or the 
atmosphere, can not be described. In addition, sources with a physically 
significant time dependencies  can not be studied, such as  supernov\ae. It 
is, however, possible to simulate beams with bunch structure, since the 
time dependence of the neutrino source is physically only important to suppress backgrounds.

On the experiment definition side, either built-in neutrino fluxes
(\eg, neutrino factory) or arbitrary fluxes can be used. Similarly,
arbitrary cross sections, energy dependent efficiencies, the
energy resolution function, the considered oscillation channels, and
many other features can be specified. For the systematics, energy
normalization and calibration errors can be simulated. Note that
the energy ranges and windows, as well as the bin widths can be
(almost) arbitrarily chosen, which means that variable bin
widths are allowed. Together with \GLOBES\ comes a number of
pre-defined experiments in order to demonstrate the capabilities
of \GLOBES\ and to provide prototypes for new experiments.

With the C-library, one can extract the $\Delta \chi^2$ for all defined 
oscillation channels for an experiment or any combination of experiments.
Of course, also low-level information, such as oscillation
probabilities or event rates, can be obtained. \GLOBES\ includes the
simulation of neutrino oscillations in matter with arbitrary matter 
density profiles, as well as it allows to simulate the matter density
uncertainty. As one of the most
advanced features of \GLOBES , it provides the technology to 
project the $\Delta \chi^2$, which is a function of all oscillation
parameters, onto any subspace of parameters by local minimization. 
This approach allows the inclusion of multi-parameter-correlations,
where external input (\eg, from solar parameters) can be imposed, too.
Applications of the projection mechanism include the projections onto the $\stheta$-axis and the $\stheta$-$\deltacp$-plane. In addition, all oscillation parameters can be kept free to precicely localize 
degenerate solutions.

\cleardoublepage
\tableofcontents

@@

As it is illustrated in \figu{GLOBES}, \GLOBES\ consists 
of several modules.
%
\begin{figure}[bht]
\begin{center}
\includegraphics[width=16cm]{GLOBES}
\end{center}
\caption{\label{fig:GLOBES} Different modules in \GLOBES .}
\end{figure}
%
\AEDL (``Abstract Experiment Definition Language'') is a language
to define experiments in form of normal text files. One or more of 
the resulting \AEDL\ files can then be processed together with supporting 
flux or cross section files by the user interface. The user interface
is a C-library, which loads one or more \AEDL\ file(s)
containing the experiment definition(s). The user interface is linked 
against the application software, which calls the user interface functions
for the intended experiment simulation.

The application 
software is, except from some example files, not part of \GLOBES , since
the evaluation of the experiment performance is often a matter of taste
and definition. In addition, the algorithms depend, especially for
high-precision instruments, very much on the oscillation parameters.
In general, it is quite simple to simulate superbeams and reactor
experiments. However, because of the more complicated topology, the
simulation of neutrino factories is much more difficult. In order
to demonstrate some of these difficulties, we show in this manual only
examples with neutrino factories. These examples can be found in
\Part~\ref{part:1} within the boxed pages. As complete files, they
are also available in the \GLOBES\ software package.

The \GLOBES\ software may have two target groups: 
Physicists, who are mainly interested in optimizing the potential
of specific experimental setups, and others, who are mainly
interested in the physics potential of different experiment types
from a theoretical point of view.
For the first group, \AEDL\ could be the most interesting aspect of
\GLOBES , where the user interface is only a tool to obtain specific
parameter sensitivities. In this case, \GLOBES\ could, serve as a
unified tool for the comparison and optimization of different experiment
 setups on equal footing, where
it is the primary objective to simulate the experiments as accurate
as possible. In addition, changes in experimental parameters, such as
efficiences or the energy resolutions, can quickly be tested.
%
For the second user group, the pre-defined 
experiment definition files might already be sufficient to test
new conceptual approaches, and the user interface is the most interesting
aspect for sophisticated applications including correlations,
degeneracies, and multi-experiment setups. In either case, the \GLOBES\
software could serve as a platform for the exchange of experiment
definitions, and for an efficient splitting of work between
experimentalists and theorists.

The user interface functions are described in \Part~\ref{part:1} of 
this manual, which is the ``user's manual''. In there, first of all a 
short \GLOBES\ tour is given in \Chapt~\ref{chapter:tour} in order to 
have an overview over \GLOBES . 
After that, the user
interface is successively introduced from very basic to more sophisticated
function. Eventually, it is demonstrated how one can change many
experiment parameters at running time (such as baseline or target mass), and how one can ontain low-level
information. We recommend that everybody interested in \GLOBES\ should
become familiar at least with the concepts in \Chapt~\ref{chapter:tour}
 and some of the examples on the boxed pages. The examples can be 
 directly compiled 
 from the respective directory in the \GLOBES\ software package.

In \Part~\ref{part:2} of the manual, \AEDL\ is described. After an
introductory chapter, all functions are defined in greater detail.
This part might be more interesting for the experimental users who
want to modify or create \AEDL\ files. A useful tool in this context
is the software program ``globes'', which returns event rates and other
information for individual \AEDL\ files without further programming. 
For example, flux normalizations can with this tool be easily adjusted 
to reproduce the event rates of a specific experiment.

Finally, \Part~\ref{part:3} of this manual serves as reference manual,
and the appendix contains further supplementary information.

@@

% Bibliography:

\bibliographystyle{./unsrt}
\bibliography{references}

% Index:

\printindex
